%
% Acknowledgements
\chapter{Preface \& Acknowledgments}%
Having a background in mechanical engineering, I have always been motivated in closing the gap between theory and application on a physical system during my master's in control systems. The topic of humanoid robotics offers a very interesting, challenging, platform to dedicate my motivation to. The complex multi-body system of a humanoid robot copes with nonlinearity, hybrid dynamics, actuation limitations and plays with your personal intuition. Also, humanoid robots are still physically far behind of what a human can do, which proves that there is large room for improvement. I believe in the future, substitution of a human with a human-like machine can be a live saving. I believe reaching out to \ac{IHMC} was the best decision to learn from and contribute to this field of research.

I would like to thank \ac{IHMC}, for giving me the opportunity to conduct research at the robotics lab. I am particularly grateful for the supervision that was given to me by Dr. Robert Griffin, Dr. Sylvain Bertrand and Dr. Jerry Pratt. I would like to thank everybody else at the robotics lab for their advice and the joy I experienced of working at the lab. 

I would also like to thank Dr. Javier Alonso-Mora from Delft University of Technology, for supervising me throughout the year I was abroad. Also, I would like to thank the examining committee members Prof. Dr. Martijn Wisse, Committee Member 4 and Committee Member 5.

I would like to thank my mother, father and two brothers, for always being supportive.

\vspace*{15mm}

Delft, University of Technology \hfill \mscname \\
\mscdate