%
\chapter{Conclusion}%
\label{chap:conclusion}
There are numerous ways to describe the dynamical behavior of a walking robot, that are often sufficient for constant walking gaits over flat terrain that are not very dynamic. Those descriptions make most of the times use of a \ac{LIP} model. If examples of human behavior while maneuvering over un-flat terrain or during rapid changes in velocity are studied, the constant height assumption is not always sufficient to describe its motion. \\
With this survey an insight has been gained in the most common humanoid robot modeling, planning and control strategies and observed that \ac{CoM} height variation of the robot is often not captured in its dynamical description.\\
A selection of one of the most relevant publications that consider height variations was reviewed, but hardware results are not often shown at robots similar to that in use at IHMC: Atlas or Valkyrie. Furthermore, often constraints were applied to system descriptions for planning and control, which have not shown yet to give better results than existing models.\\
The discussed methods that link to \ac{CoM} height variations can be broken down based on:
\begin{itemize}
\item \ac{2D} and \ac{3D} methods.
\item Methods that need a predefined footstep plan and timing and methods that capture this in the problem.
\item Computation times.
\item Theory or application orientated methods.
\item Robustness against disturbances, in the case of a control method.
\end{itemize}
Concerning the goal of application, \ac{2D} side-view methods, like \cite{pratt2007derivation, koolen2016balance, gao2017increase, nguyen2017dynamic}, need special care compared to \ac{3D} methods, like \cite{englsberger2013three, hopkins2014humanoid, liu2015trajectory, caron2018capturability}. Methods that need a predefined footstep plan and timing, like  \cite{englsberger2013three, hopkins2014humanoid,  caron2018capturability} have an advantage over those who do not, as most humanoid robots rely on a footstep plan. Strategies with fast computation times, like the analytic \ac{MPC} approach in \cite{koolen2016balance}, are more suited for real-time usage. Robustness properties are an important factor in control, but are not often shown.\\
Regarding the exploration of the effects of height variation, all different insights have a contribution to this. Fundamental theory can be explanatory and might be useful to extend to application.  \\
This literature survey emphasizes the idea that it is an interesting challenge to investigate the effects of height variation in the humanoid robot model and to see to what extend this can be applied on a physical robot. The goal in the research project will be to develop new varying height models and to build on existing models, to extend current knowledge and to find an application on hardware that improves current planning and control techniques.\\
The final implementation may have the more complicated shape of a trajectory planning or a \ac{MPC} problem, or may use simple heuristics that are implemented in a smart way. Work can be done on for example extending the \ac{Eorbit} of \cite{pratt2007derivation} to a \ac{3D} version, so this becomes applicable on hardware. In a real-time planning or control problem, computation time is an important aspect, and different strategies have to be judged on speed as well. 